Increased perceptions of autonomy through choice fail to enhance motor skill retention

### 2021-11-04 docs/r1_submission/r1_responses-to-reviewers.Rmd

The measured related to intrinsic motivation used in the present experiment is consistent with that used in previous experiments (e.g., Barros et al. 2019). While Sanli et al. (2013) and Wulf and Lewthwaite (2016) discuss how perceived autonomy through choice can enhance intrinsic motivation and better motor learning, Ste-Marie et al. (2015) ran a path analysis and showed that intrinsic motivation is not predictive of the self-controlled learning advantage. We elected to not run an exploratory analysis to test whether intrinsic motivation predicted post-test performance, regardless of group. We believe this not only detracts from the main question of our experiment, but there is not prediction in the "OPTIMAL" theory surrounding intrinsic motivation predicting post-test performance, independent of group. As mentioned in one of our earlier responses, we have made specific reference to the relevant predictions in the "OPTIMAL" theory (see Pages 4-5, Lines 18-27 and 1-4).

@@

# Reviewer \#3

\RC{\emph{It was a pleasure to read and review your article "Increased perceptions of autonomy through choice fail to enhance motor skill retention" submitted to the Journal of Experimental Psychology: Human Perception and Performance. Your writing is apparent and concise. This article examines autonomy-support during the speed cup-stacking task. There were three conditions: the self-control group, the traditional yoked group, and the explicit yoked group. The results suggest that there are no statistically significant learning differences between the groups.
}}

Thank you for the positive comments about our manuscript.

\RC{\emph{Public significance statement, Lines 21-23, what kind of effective practice variable should coaches and practitioners use? Autonomy support affects skill learning among various populations. For instance: (https://www.tandfonline.com/doi/abs/10.1080/17461391.2020.1720829)
}}

Based on your comment and some reflection/discussions among ourselves, we decided to modify our public significance statement to the following:

> The present study suggests that although university-aged students may experience a boost in feelings of autonomy when given choice over features of their learning environment, this boost may not enhance learning of a new motor skill. Our results also highlight the need for higher powered and pre-registered motor learning experiments if coaches and practictioners are to be provided with reliable recommendations for training and rehabilitation protocols.

\RC{\emph{Equivalence tests? Why is it important? I have little experience with equivalence tests.
}}

When a null hypothesis significance test returns a non-significant result, many researchers incorrectly conclude that there is no effect or that groups are not different. Equivalence tests provide a statistical tool that researchers can use to support the absence of a meaningful effect. In other words, an equivalence test can help researchers conclude that a non-zero difference between groups is practically equivalent to zero.

\RC{\emph{Page 4, lines 17-20, does autonomy support differ between adults and children? Can you explain the findings of this study?
}}

The reference here is to the paper that introduces the "OPTIMAL" theory of motor learning, which argues that the learning benefits of self-controlled practice conditions are "very robust and generalizable to different tasks, age groups, and populations (p. 1393).

\RC{\emph{Lines 23-24 review the sentence.
}}

This sentence has been removed in the revised manuscript based on comments from Reviewer 2.

\RC{\emph{In lines 25-27 you mention previous literature does not indicate that self-controlled practice is autonomy-supportive, but in lines 17-18, p.4, you say a variety of names and labeled them as autonomy-supportive. Could you clarify this?
}}

We have updated our wording in these lines to clarify our position as our intention was not to label these names as being autonomy-supportive. This part now reads as (see Page 4, Lines 16-21):

> Over the years, this manipulation has been described using a variety of names (e.g., self- and/or learner-controlled; -regulated; -directed; -selected), but more recently some researchers have adopted the term autonomy-support. Within their “OPTIMAL” theory of motor learning, Wulf and Lewthwaite (2016) argued that providing learners with opportunities for choice creates an autonomy-supportive practice environment, which facilitates motor performance and learning.

\RC{\emph{ What was your hypothesis or prediction? Did your results support the hypothesis?
}}

We have added this information (see Pages 7-8, Lines 22-27 and 1-7):

> We predicted that the self-controlled group would have significantly faster stacking times in retention than the traditional yoked group. In other words, we expected to replicate the typical self-controlled learning advantage. If the self-controlled learning advantage results from choice being autonomy-supportive as argued in the “OPTIMAL” theory (Wulf & Lewthwaite, 2016), then participants in the self-controlled group should also self-report significantly higher scores for perceptions of autonomy. Also based on the view that autonomy-support is the mechanism underlying the typical self-controlled learning advantage (Wulf & Lewthwaite, 2016), we predicted that the explicit yoked group should have significantly slower stacking times in retention. Although we found significantly higher perceptions of autonomy in the self-controlled group, there were no significant group differences in stacking times during retention. Overall, these results do not support our predictions and are inconsistent with predictions 2, 3, and 7 in the “OPTIMAL” theory of motor learning (Wulf & Lewthwaite, 2016).

\RC{\emph{Thank you for providing additional material at asPredicted.org, it was very beneficial.
}}

You're welcome.

\RC{\emph{Lines 18-19, p.7, what age are the participants? Where were they recruited from? Who are we looking at? (e.g., university students, adults, etc.)?
}}

We have made specific reference that the participants were university students at the start of the Participants section (see Page 8, Line 15). It seems like the reported mean age of the participants in each group was missed. This information is also reported in the Participants section.

\RC{\emph{There was only one analysis explained, it would help if you discussed all analyses. Explain the analysis you ran in your data analysis section. 
}}

We described all analyses in the Results section as a way to reduce the cognitive load placed on the reader. We believe this prevents the reader from having to either remember all statistical analyses by the time they reach the Results section or having to scroll back and forth between the Data analysis section and the Results section. We have added a line to the Data analysis section that informs the reader that statistical tests are described below (see Page 12, Line 9).

\RC{\emph{Why do you have other graphs? You did not run an analysis on each trial but ran three by five.
}}

We have additional figures for transparency and for the interested reader. Figures 1A and 1D further support the pattern of results in Figures 1B and 1C and the associated outcomes of the statistical tests.

\RC{\emph{p.12, figure 1C, adjusted time is not explained in the results; what is adjusted mean stacking? Some of the graphs can be removed from the paper.
}}

Adjusted mean stacking time is stacking time in the retention test, adjusted for pre-test scores. We think our explanation for this may have been missed. In the Results section, we stated that "means are shown as the adjusted means controlling for pre-test" and the figure caption for 1C is "Mean stacking time (s) adjusted for pre-test for each group". We have elected to not removed any figures based on our response in the above comment.

\RC{\emph{Lines 10-17, p.13, please indicate the figure the reader should reference with your results. What graph is to be looked at when you ran your analysis? 
}}

We think our reference to which figures to look may have been overlooked as these were included in the relevant subsections of the Results section.

\RC{\emph{p.13, why did you use a Welch's t-test to compare two groups? Why not run a Welch's ANOVA between the three experimental groups? You had three experimental groups but only analyzed two, why?
}}

The analysis involving all three groups was run in our pre-registered analysis. As mentioned in the subsection Traditional self-controlled learning advantage, testing for a the typical self-controlled learning advantage involves a comparison between the Self-Controlled group and the Traditional Yoked group. This is why a Welch's t-test was used to test for this non-preregistered analysis.

\RC{\emph{p.16, figure 2, I don't understand why you did this instead of bar graphs. To highlight group differences, why not use bar graphs instead of having every single data point on it. If no self-control effect was found, how can we trust the questionnaire data? 
}}

Bar graphs have been criticized as many different datasets can produce the same bar graph (see Figure 1 in Weissgerber et al. 2015, https://doi.org/10.1371/journal.pbio.1002128). For this reason, we show the individual data points in addition to the group mean for greater transparency.

This is a reasonable question; however, we are not sure a self-controlled learning advantage is required to trust questionnaire data. Despite the claim that the provision of choice is a very robust and generalizable learning variable in the "OPTIMAL" theory, the recent meta-analysis of the self-controlled literature we highlight in the discussion strongly challenges this assertion (McKay et al. Accepted). Thus, the provision of choice may only exert an influence on perceptions of autonomy (if you have a large enough sample size to detect this small effect), which in itself may be a desired outcome in some situations.

\RC{\emph{Can you explain why you have pre-registered and registered analyses separate? Why is a one-way ANCOVA the only pre-registered analysis?
}}

The ANCOVA is our only pre-registered analysis as this was our first time doing a pre-registration and we overlooked specifying the other analyses typically run in a self-controlled motor learning experiment. It is for this reason, we separated our analyses into a Pre-registered analysis and Non pre-registered analyses to be transparent that only one analysis was pre-registered.

\RC{\emph{Can you address the equivalence tests? Most motor learning researchers are not familiar with these tests, why are equivalence tests used and needed in this study? 
}}

We have added a description for why we included equivalance tests in the present experiment (see Page 16, Lines 7-13):

> Given our pre-registered analysis resulted in a non-significant finding, we performed two equivalence tests (Self-Controlled versus Traditional Yoked and Traditional Yoked versus Explicit Yoked) on pre-test adjusted retention stacking times using the two one-sided test procedure (Lakens et al., 2018; Schuirmann, 1987). Equivalence tests are a statistical tool that researchers can use to support the absence of a meaningful effect and avoid incorrectly concluded no effect exists based on a non-significant finding (Harms & Lakens, 2018).

\RC{\emph{How many papers use cup stacking? How do you know that it is not the task itself? Could the issue be with the task you chose? Why not use a task that is commonly used in motor learning? Is it an issue with the task, or is it an issue with the manipulation?
In summary, this research is interesting and well done. When major revisions are addressed, the authors should take advantage of the interesting information available to them and try to publish this work.
}}

We did a quick google scholar search for "motor learning cup stacking", which resulted in over 6500 papers. This is in line with our perception that cup-stacking is a commonly used task in motor learning research. Thus, we do not think a task argument is valid for the lack of a self-controlled learning advantage. We think the underlying issue is that the view within the "OPTIMAL" theory that self-controlled practice conditions are a very robust and generalizable learning variable is simply wrong as supported by the trivial effects found in the meta-analysis by McKay and colleagues. As mentioned in the Discussion section, our estimated effect for self-controlled versus yoked fell within the plausible range of these trivial effects.

<!-- \clearpage -->

<!-- # References -->

<!-- \begingroup -->

<!-- <div id="refs"></div> -->

<!-- \endgroup -->
